
# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

### 1. Load the data (i.e. read.csv())

```{r data_prepro, echo=TRUE}
setwd("C:/Users/mmahajan/Desktop/Temp Files/Kag/R")
activityData <- read.csv("C:/Users/mmahajan/Desktop/Temp Files/Kag/R/activity.csv", header=T, sep=",")
```
### 2. What is mean total number of steps taken per day?


```{r mean, echo=TRUE}
stepsByDay <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)

```
### 2. Calculate and report the mean and median total number of steps taken per day

```{r mean_med, echo=TRUE}
hist(stepsByDay,data = stepsByDay)

```

## Imputing missing values

```{r ggplot12, echo=TRUE}
 
numMissingValues <- length(which(is.na(activityData$steps)))

numMissingValues
activityDataImputed <- activityData
library(Hmisc)
activityDataImputed$steps <- impute(activityData$steps, fun=mean)
stepsByDayImputed <- tapply(activityDataImputed$steps, activityDataImputed$date, sum)


hist(stepsByDayImputed,data = stepsByDayImputed)


stepsByDayMeanImputed <- mean(stepsByDayImputed)
stepsByDayMedianImputed <- median(stepsByDayImputed)

 stepsByDayMeanImputed
 stepsByDayMedianImputed

```

## Are there differences in activity patterns between weekdays and weekends?

```{r ggplot13, echo=TRUE}
activityDataImputed$dateType <-  ifelse(as.POSIXlt(activityDataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')

averagedActivityDataImputed <- aggregate(steps ~ interval + dateType, data=activityDataImputed, mean)

plot(averagedActivityDataImputed$interval,averagedActivityDataImputed$steps,type='l', col='red')
  
```



### 3. Calculate and report the mean and median total number of steps taken per day
```{r hist, echo=TRUE}
stepsByDayMean <- mean(stepsByDay)
stepsByDayMedian <- median(stepsByDay)
stepsByDayMean
stepsByDayMedian

```


## What is the average daily activity pattern?

```{r ggplot11, echo=TRUE}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=activityData$steps), by=list(interval=activityData$interval), FUN=mean, na.rm=TRUE)


plot(averageStepsPerTimeBlock$interval,averageStepsPerTimeBlock$meanSteps,type='l', col='red')


mostSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[mostSteps,'interval'])

timeMostSteps  

```

## Imputing missing values

```{r ggplot21, echo=TRUE}
 
numMissingValues <- length(which(is.na(activityData$steps)))

numMissingValues
activityDataImputed <- activityData
library(Hmisc)
activityDataImputed$steps <- impute(activityData$steps, fun=mean)
stepsByDayImputed <- tapply(activityDataImputed$steps, activityDataImputed$date, sum)

hist(stepsByDayImputed,data = stepsByDayImputed)


stepsByDayMeanImputed <- mean(stepsByDayImputed)
stepsByDayMedianImputed <- median(stepsByDayImputed)

stepsByDayMeanImputed
stepsByDayMedianImputed


```

## Are there differences in activity patterns between weekdays and weekends?

```{r ggplot31, echo=TRUE}
activityDataImputed$dateType <-  ifelse(as.POSIXlt(activityDataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')

averagedActivityDataImputed <- aggregate(steps ~ interval + dateType, data=activityDataImputed, mean)

plot(averagedActivityDataImputed$interval,averagedActivityDataImputed$steps,type='l', col='red')
  
```





